\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-29 chapters/gaze_independence.tex
Layout conclusion
Second layout run
Accuracy figure

\chapter{Gaze-independent visual \acsp{bci}}
\label{sec:gaze-independent}
\epigraph{%
  %\justifying
  ``\elide\ This begs the question: if a patient had reliable control
  over their eye gaze, why would one aim to facilitate communication with that
  patient via a BCI? It would be far simpler and more efficient to use an
  eye-tracking system.
  And even if a BCI were the preferred or only available option for some reason,
  designs requiring eye gaze control could not be used
  by patients with severe motor impairment, as these patients
  cannot direct their gaze.
  Because these are the patients who have the greatest need for BCIs, it is
  important that accurate gaze-independent BCIs exist.''
}{%
  \textcite{Egan2017}
}

@@

\newcommand\fnwolpawcrit{\footnote{
``The first class consists of people who are truly totally locked-in (e.g.,
due to end-stage ALS or severe cerebral palsy), who have no remaining
useful neuromuscular control of any sort, including no eye movement.
\elide\ This class is very small. \elide\
The second class of potential \ac{bci} users comprises those who retain
a very limited capacity for neuromuscular control. This group includes
people who retain some useful eye movement or enough limb muscle
function to operate a single-switch system. Such control is often slow,
unreliable, or easily fatigued.
This group is much larger than the first.''~\cite{Wolpaw2006}}}

@@

\begin{figure}
  \input{figures/gaze_independence/lis_venn.tikz.tex}
  \caption[Venn diagram of potential users of \acs{bci} assistive technology.]{%
    Venn diagram of potential users of assistive technology.
    Patients with \acf{sspi} are the general target population.
    If eye gaze-based solutions are not properly suited for an individual,
    they have \acf{sspgi}.
    Individuals with \ac{lis} have no muscular control left, or it is very limited,
    such as the ability to communicate through a binary switching system.
    \Acp{bci} are especially relevant for the last two groups, as they are not
    able to use eye tracking solutions.
  }
  \label{fig:gaze-independence/venn}
\end{figure}

>>>>

\chapter{Gaze-independent visual \acsp{bci}}
\label{sec:gaze-independent}
\epigraph{%
  \justifying
  ``\elide\ This begs the question: if a patient had reliable control
  over their eye gaze, why would one aim to facilitate communication with that
  patient via a BCI? It would be far simpler and more efficient to use an
  eye-tracking system.
  And even if a BCI were the preferred or only available option for some reason,
  designs requiring eye gaze control could not be used
  by patients with severe motor impairment, as these patients
  cannot direct their gaze.
  Because these are the patients who have the greatest need for BCIs, it is
  important that accurate gaze-independent BCIs exist.''
}{%
  \textcite{Egan2017}
}

@@

\newcommand\fnwolpawcrit{\footnote{
``The first class consists of people who are truly totally locked-in (e.g.,
due to end-stage ALS or severe cerebral palsy), who have no remaining
useful neuromuscular control of any sort, including no eye movement.
\elide\ This class is very small. \elide\
The second class of potential \ac{bci} users comprises those who retain
a very limited capacity for neuromuscular control. This group includes
people who retain some useful eye movement or enough limb muscle
function to operate a single-switch system. Such control is often slow,
unreliable, or easily fatigued.
This group is much larger than the
first.''~\cite{Wolpaw2006}\label{fn:bci/wolpawcrit}}}

@@

\begin{figure}[t]
  \input{figures/gaze_independence/lis_venn.tikz.tex}
  \caption[Venn diagram of potential users of \acs{bci} assistive technology.]{%
    Venn diagram of potential users of assistive technology.
    Patients with \acf{sspi} are the general target population.
    If eye gaze-based solutions are not properly suited for an individual,
    they have \acf{sspgi}.
    Individuals with \ac{lis} have no muscular control left, or it is very limited,
    such as the ability to communicate through a binary switching system.
    \Acp{bci} are especially relevant for the last two groups, as they are not
    able to use eye tracking solutions.
  }
  \label{fig:gaze-independence/venn}
\end{figure}

### 2024-11-29 cover_kul.tex
All review comments
Accuracy figure

\begin{titlepage}
    \sffamily
    \begin{minipage}[t]{.5\textwidth}
    \vspace{0pt}
    KU Leuven \\
    Biomedical Sciences Group \\
    Doctoral School of Biomedical Sciences \\
    Faculty of Medical Science \\
    Department of Neurosciences
    \bigskip

@@

    \noindent\makebox[\textwidth][c]{%
      \begin{minipage}{.8\textwidth}
        \centering
        {%
          \Huge
          \bfseries
          A visual Brain-Computer Interface for gaze-free communication
        }
        \bigskip

@@

    \vfill
    \begin{minipage}[b]{.5\textwidth}
      \raggedright
      \small
      Jury: \\
      Co-supervisor KU Leuven: \emph{prof. Marc Van Hulle} \\
      Co-supervisor Univ. Lille: \emph{prof. François Cabestaing} \\
      Chair examining committee: \emph{prof. John Creemers} \\
      Chair public defense: \emph{prof. Koen Poesen} \\
      \emph{prof. Andrea K\"ubler} \\
      \emph{prof.\ Fabien Lotte} \\
      \emph{prof.\ Reinhold Scherer} \\
      \emph{prof.\ Maarten De Vos} \\
      \emph{prof.\ Adalberto Simeone}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.31\textwidth}
        \raggedright
        \small
        Dissertation presented in partial fulfillment of the requirements for
        the double degree of \emph{Doctor of Biomedical Sciences} (KU Leuven)
        and \emph{Doctor of Control Science and Signal Processing} (University
        of Lille)
    \end{minipage}
    \bigskip

    \centering
    Defended December 16, 2024

\end{titlepage}

>>>>

{%
\small
\parindent0pt
  \sffamily
  \begin{minipage}[t]{.5\textwidth}
  \vspace{0pt}
  KU Leuven \\
  Biomedical Sciences Group \\
  Doctoral School of Biomedical Sciences \\
  Faculty of Medical Science \\
  Department of Neurosciences
  \bigskip

@@

  \makebox[\textwidth][c]{%
    \begin{minipage}{\textwidth}
      \centering
      {%
        \huge
        \bfseries
        A visual Brain-Computer Interface for gaze-free communication
      }
      \bigskip

@@

  \vfill
  \begin{minipage}[b]{.6\textwidth}
    \raggedright
    Jury: \\
    Co-supervisor KU Leuven: \emph{prof. Marc Van Hulle} \\
    Co-supervisor Univ. Lille: \emph{prof. François Cabestaing} \\
    Chair examining committee: \emph{prof. John Creemers} \\
    Chair public defense: \emph{prof. Koen Poesen} \\
    \emph{prof. Andrea K\"ubler} \\
    \emph{prof.\ Fabien Lotte} \\
    \emph{prof.\ Reinhold Scherer} \\
    \emph{prof.\ Maarten De Vos} \\
    \emph{prof.\ Adalberto Simeone}
  \end{minipage}\hfill%
  \begin{minipage}[b]{.31\textwidth}
      \raggedright
      Dissertation presented in partial fulfillment of the requirements for
      the double degree of \emph{Doctor of Biomedical Sciences} (KU Leuven)
      and \emph{Doctor of Control Science and Signal Processing} (University
      of Lille)
  \end{minipage}
  \bigskip

  \centering
  \hfill Defended December 16, 2024\hfill
}

### 2024-11-29 cover_ulille.tex
Accuracy figure

{%
\parindent0pt
    \sffamily
    \begin{minipage}[t]{.5\textwidth}
    \vspace{0pt}
    Universit\'e de Lille \\
    \'Ecole Doctorale MADIS \\
    UMR CRIStAL
    \bigskip

@@

    \vfill
      \small
      Th\`ese soutenue le 16 d\'ecembre 2024 en satisfaction partielle des exigences de double
      dipl\^ome de Docteur en Traitement du Signal et des Images
      (Universit\'e de Lille) et Docteur en Sciences Biom\'edicales (KU Leuven)
      devant le jury compos\'e de: \\
      \smallskip

>>>>

{%
  \small
\parindent0pt
    \sffamily
    \begin{minipage}[t]{.5\textwidth}
    \vspace{0pt}
    Universit\'e de Lille \\
    \'Ecole Doctorale MADIS \\
    UMR CRIStAL
    \bigskip

@@

    \vfill
      Th\`ese soutenue le 16 d\'ecembre 2024 en satisfaction partielle des exigences de double
      dipl\^ome de Docteur en Traitement du Signal et des Images
      (Universit\'e de Lille) et Docteur en Sciences Biom\'edicales (KU Leuven)
      devant le jury compos\'e de: \\
      \smallskip
